\documentclass[times]{elsarticle}
\usepackage{fullpage}
\usepackage{xcolor}
\usepackage{natbib}
\usepackage[hidelinks]{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[final,babel]{microtype}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage{csquotes}
\usepackage[T1]{fontenc}
\usepackage{enumitem}

\newcommand{\TODO}[1]{TODO: \textit{#1}}
\newcommand{\revone}[1]{\textcolor{purple}{\textbf{#1}}}
\newcommand{\revtwo}[1]{\textcolor{teal}{\textbf{#1}}}
\newcommand{\revthree}[1]{\textcolor{olive}{\textbf{#1}}}
\newcommand{\revother}[1]{\textcolor{orange}{\textbf{#1}}}

\newlist{comment}{enumerate}{1}                                                 
\setlist[comment,1]{label={\arabic{commenti}.},leftmargin=*,resume=comment}

\begin{document}
\begin{frontmatter}
\title{Responses to reviewers \\
\vspace*{0.5em}
\large{A multidimensional method-of-lines transport scheme for atmospheric flows over steep terrain using arbitrary meshes} \\
\vspace*{0.5em}
\normalsize{JCOMP-D-17-00161}}
\author{James Shaw}
\author{Hilary Weller}
\author{John Methven}
\author{Terry Davies}
\end{frontmatter}

We greatly appreciate the reviewers' detailed comments and questions.  The revised manuscript highlights changes with a bold typeface and the text colour corresponds to the section headings below.  The revised manuscript also highlights a small number of additional changes \revother{in orange}.

\section*{\revone{Reviewer \#1}}
This paper presents a multidimensional method-of-lines finite-volume transport scheme,
which is second-order accurate and offers high solution quality and robustness under
strongly deformed meshes. Among others, the latter is of increasing importance for 
atmospheric models with complex orography. The proposed scheme is named "cubic-fit", 
as it fits multidimensional polynomials of up to third-order over an upwind stencil
using a least-squares approach. The construction of the polynomials takes into account 
constraints derived from a linear stability analysis for stencils in regions of  
highly-deformed computational meshes.

The accuracy and stability of the cubic-fit scheme is studied with a number of well 
chosen test cases that are relevant to atmospheric models. 

In my view, the presented developments definitely warrant publication in JCP. 
The content of the paper appears of high quality and adheres to the JCP standards. 
The text is well structured and clearly written. It only needs little improvement 
as given by the subsequent comments. 

I recommend acceptance of the paper subject to minor revisions by the authors.

\subsection*{Comments}

\begin{quotation}
\begin{comment}
\item  It is explained in various places that the cubic-fit scheme
    offers favourable computational efficiency due to the fact
    that most of the reconstruction depends on the mesh geometry
    only, and this part can be precomputed. It would be very interesting
    to see numbers in the paper. It doesn't have to be a comprehensive
    study, but just to get an impression of the cost, e.g. how does
    the runtime of the cubic-fit compare to the second-order upwind
    scheme? Or, if one runs cubic-fit and the upwind scheme such that
    both produce about the same $\ell_2$ error (by using a higher resolution
    with the upwind scheme), what is then the difference in runtime?
\end{comment}
\end{quotation}
\TODO{}

\begin{quotation}
\begin{comment}
\item  As only 2D simulations have been considered in the paper, I
    wondered how difficult it will be to extend the cubic-fit
    scheme to 3D? It could useful to provide a short comment
    about this, just whether complications are anticipated or
    not. (See also the subsequent comment.)
\end{comment}
\end{quotation}
\TODO{}

\begin{quotation}
\begin{comment}
\item  Appendix A, p.22: How restrictive is the 1D stability analysis
    for the multidimensional domains? May this become a more
    serious issue in 3D configurations?
\end{comment}
\end{quotation}
\TODO{}

\begin{quotation}
\begin{comment}
\item  In various places throughout the paper "icosahedra" is used
    instead of "icosadedral". I suggest the latter should be used
    consistently throughout the paper.
\end{comment}
\end{quotation}
\TODO{}

\begin{quotation}
\begin{comment}
\item  p.3, end of first paragraph: I suggest to mention here a
    recent advancement of the finite-volume MPDATA (Kühnlein and
    Smolarkiewicz, J. Comput. Phys. 2017,
    \url{http://dx.doi.org/10.1016/j.jcp.2016.12.054}). Among others, this
    paper demonstrates applicability of MPDATA for 3D compressible
    atmospheric dynamics on arbitrary hybrid unstructured meshes.
\end{comment}
\end{quotation}
\TODO{}

\begin{quotation}
\begin{comment}
\item  p.3, paragraph after Eq. (2b): if there is no specific need, I
    sugggest to remove "zero-dimensional".
\end{comment}
\end{quotation}
\TODO{Hilary?}

\begin{quotation}
\begin{comment}
\item  p.4, last paragraph in Section 2.0: "methods are described next"
    instead of "methods described next".
\end{comment}
\end{quotation}
Fixed.

\begin{quotation}
\begin{comment}
\item  p.5, first paragraph: Is it "arbitrary meshes" or "arbitrary
    structured meshes"?
\end{comment}
\end{quotation}
\TODO{}

\begin{quotation}
\begin{comment}
\item  p.5, last paragraph of Section 2.1.1.: "dirichlet" should be
    upper case "Dirichlet".
\end{comment}
\end{quotation}
Fixed.

10.) p.7, caption of Fig. 3, last sentence: It is stated that a
    von Neumann boundary condition is assumed. However, the
    footnote on p.5 says that Neumann BCs are excluded from the
    set of stencil boundary faces?

11.) p.9, caption of Fig. 4, first sentence: "A one-dimensional
    least squares fits" should probably read "One-dimensional
    least squares fits".

12.) p.12, caption of Fig. 6, last but one sentence: "domain is
    shown" instead of "domain in shown".

13.) Section 3: Why are most of the experiments run with maximum CFL
    numbers < 0.5 or even < 0.4, although the cubic-fit scheme
    permits a maximum CFL < 1 (Fig. 10)? Is this because the upwind
    scheme, to which cubic-fit is compared to, is more limited with
    respect to the time step for the considered test cases?

14.) p. 18, last but one paragraph, last sentence: correct "imprintingin".



\section*{\revtwo{Reviewer \#2}}


\subsection*{General remarks}

The manuscript presents a new method-of-lines based transport scheme
in which fluxes are computed by fitting a multidimensional
polynomial over an upwind-biased set of grid cells. The scheme
has low computation and memory costs. It is suitable for
structured and unstructured grids, and maintains accuracy
on strongly distorted grids such as those used in atmospheric models
over steep terrain. An idealized analysis suggests how the
construction of the polynomial should be constrained to obtain
conditional stability of the scheme, and such stability appears
to be achieved in practice.
I believe the manuscript will be of interest to those developing
atmospheric models and to the wider CFD community. I would be happy to
recommend publication after the points below are addressed.


\subsection*{Main points}

1. It is claimed (in several places) that the scheme requires `just one
vector multiply per face per time step'. Since the two-stage Heun scheme
(2) is used, surely the cost must be two vector multiplies per face per time
step (or, more generally, one vector multiply per face per {\it stage})?

2. It is claimed in several places that the scheme is stable, but some qualification
is appropriate. First, the scheme is {\it conditionally} stable; the Courant number
should not be too large. Then Appendix~A gives an analysis deriving constraints
on the polynomial reconstruction that should be satisfied for stability. However,
the analysis is idealized in various ways: one-dimensional, regular grid, constant $v$,
two- and three-cell stencils, and ignoring the time discretization (also see below).
That is fair enough---it might not be possible to generalize the analysis to the full
scheme---and it is clearly stated at the end of the Appendix that the constraints are
{\it hypothesized} to apply more generally. However, in section 2.1.4 the language is
stronger and suggests that these constraints are definitive: ``...stability constraints...'',
``...must satisfy...'', etc.

3. P14. The calculation of the longest stable timestep {\it assumes} that the stability
limit is given by $\mathrm{Co}=1$. Given the idealized nature of the analysis in
Appendix~A, and that most experiments are done with $\mathrm{Co} \approx 0.4$,
at least some empirical investigation of the stability limit should be discussed
to confirm that it is indeed (exactly or approximately) $\mathrm{Co}=1$.

4. The presentation of the analysis in Appendix~A is somewhat confusing. It appears
to use a time discretization by introducing a time step $\Delta t$ and an amplification
factor $A$, (and hence a Courant number $v \Delta t / \Delta x$) whereas the
analysis is, in fact, continuous in time. It might be clearer to set
$\phi(x_j , t) = \hat{\phi}(t) \mathrm{e}^{i j k \Delta x}$ and form an ODE for
$\hat{\phi}$, etc. I think the conclusion would be the same.

Might it be possible to include the time discretization (2) in the stability analysis?
It would be interesting to know if the weak instability of (2) (noted on p3) affects
the constraints (57), (58), (63).

5. The Courant number as defined by (4) would appear to vanish for a non-divergent
flow. A more natural definition might be to sum only over outflow faces (and
remove the factor $1/2$). For non-divergent flow that would be equivalent to
replacing $\mathbf{u} \cdot \mathbf{S}_f$ by $| \mathbf{u} \cdot \mathbf{S}_f |$
in~(4).

6. Section 3.2 and Table~1: what resolution is used?

\subsection*{Minor points}

It would be interesting to know if there was some rationale for the choice of
the polynomial (5) and the stencil discussed in  section 2.1.1. Given that the scheme
is second order convergent, might it be possible to use a lower degree polynomial
and smaller stencil?

\vspace{2mm}

The first two paragraphs on p2 refer to quadrilateral meshes. It seems the discussion
has slipped into the 2D vertical slice case without explicitly saying so.

\vspace{2mm}

P2, 5th paragraph. Perhaps mention that dimensionally-split schemes are only possible
when the grid permits it (e.g.\ not a hexagonal grid).

\vspace{2mm}

P9, last sentence of section 2.1. Should $w_d$ be $m_d$?

\vspace{2mm}

The discussion of previous work is generally very good. On p2 is it worth mentioning
the SLICE scheme of Zerroukat and co-authors? In the first paragraph of section 3.3, credit
for the hexagonal-icosahedral mesh should probably be given to Heikes and Randall.

\vspace{2mm}

Some brief comments on how the approach extends to three dimensions would be
appropriate.

\vspace{4mm}

\subsection*{Typos, etc}

Introduction, second sentence: `modification of the lower boundary' is an odd way to
think of it.

P2, 5 lines from the bottom: `...non-simply connected domains...'. I found this
phrase confusing.

P3 line 17: `cubic, upwind-biased stencil'. Surely it is the polynomial that is
cubic, rather than the stencil?

P3 below (2): `stable ... with some upwinding'. Presumably some qualification is required
(with sufficient upwinding)?

P5 footnote: could be made

P14 5 lines from the bottom: cubic should be cubicFit?

P18 2 lines from the bottom: imprinting in


\section*{\revthree{Reviewer \#3}}
The author describes  a standard finite volume algorithm for solving the advection equation on unstructured
grids. Using a method of lines the spatial discretization consists in computing face values of a cell
centered given function by high order polynomial point wise approximation in contrast to a more common
area wise procedure. The proposed algorithm has three phases, choose of two upwind dominated stencils
for each face, determination of a subset of candidate polynomials with at most order three, weighting
of the individual contributions in a least squares approximation to increase stability with respect 
to the advection equation.

The stencil choice starts with the determination of opposing faces in the upwind cell with respect to 
to the destination face, and adding further cells in a two step procedure. The candidate polynomials
with at most order three are all polynomials where the corresponding least squares problem has full rank.
For the third step the candidate polynomials are ordered with respect to the number of unknowns and 
and larger minimal singular value in case of polynomials with the same number. In the second step
following the first ordering a weighting sequence is tested until a stability criterion is fulfilled
which can finally end in a simple upwind scheme. 

The paper is well written and illustrates all steps by intuitive examples. The new test example
for the two-dimensional example with the boundary following velocity field enriches the number of
problems for testing numerical advection schemes.

\subsection*{Comments and questions}

Can you comment on the stencil for triangular grids. In case of a slightly perturbed triangular grid
with  equilateral triangles the stencil changes in comparison the unperturbed case.

Why do you prefer the point wise approximation?

The singular value test should be relative to the largest singular value.

The weighting can degrade the condition number of the least squares problem.

Is there a smoothing of  hexagonal-icosahedral mesh after refinement.

Can you comment on the projection to the tangent plane for the spherical test example. 
Note that projection to another grid can change the interpolation/approximation essentially, 
like the transformation of a boundary following grid in physical space to a Cartesian grid in 
computational space. Why is the interpolation/approximation not done in 3 dimensional 
physical space.

\bibliographystyle{elsarticle-num}
\bibliography{references}
\end{document}
